### 2022-09-16 syntax.tex
All the files TODOs separated

\section{Syntax}
\label{sec:syntax}
\mytodo{Ismail's Part}