### 2022-09-17 syntax.tex
separted stuff

\section{Syntax: Sequencing \& Executing Instructions}
\label{sec:syntax}
In this paper, we introduce a simple language for createing a stream of instructions and executing a single one to change the machine state.
\input{fig-syntax}